The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-02-20 src/ms.tex
Some comments on numpy and units

  The modular nature of \astropypkg means that not all sub-packages deeply
  integrate with one another -- and similarly between separate \python
  libraries. Because \astropy is situated at the nexus between scientific
  computing into astronomy, interoperability has been and continues to be an
  important focus for \astropy, both within the various \astropypkg sub-packages
  and with other important \python libraries. While increasing inter and
  intra-package integration will be a perpetual goal -- with feature additions
  in one place requiring updates in another -- \astropy has made impactful
  progress with units, tables, and I/O.
  % Better list? Also better transition
  If done right, all a user notices is that code using many different python
  features and libraries ``just works''. Here we will draw notice to these
  efforts to make your code ``just work''.

@@

    \astropyQuantity\ is a backbone of \astropy, leveraging the power of
    \package{numpy} and adding units. Previously, many \package{numpy} functions
    would strip a \astropyQuantity\ of its units (or fail outright), limiting
    \astropyQuantity's usefulness. Now, advances in \package{numpy} function
    overloading (see for example
    \href{https://numpy.org/neps/nep-0018-array-function-protocol.html}{NEP 18})
    mean \astropyQuantity\ works with almost all \package{numpy} (version 1.17+)
    functions. These advances have similarly benefitted many
    \package{numpy}-dependent packages, like \package{scipy} and their sparse
    matrices. And again \astropypkg benefits since more \package{scipy}
    functions now work with \astropyQuantity.

    In the remaining gaps, the \package{numpy} and \astropyQuantity\
    interoperability efforts are ongoing. For some functions, such as in
    \href{https://numpy.org/doc/stable/user/basics.rec.html#module-numpy.lib.recfunctions}{\package{numpy}'s
    module} for manipulating structured arrays, compatibility only requires
    exending the existing \package{numpy}-\astropypkg bridge framework.
    Community interest, in the form of a Feature Request (or better yet Pull
    Request) would be sufficient to see this realized. For other functions,
    discussed further in the
    \href{https://docs.astropy.org/en/stable/known_issues.html#known-deficiencies}{\astropy
    documentation}, the \package{numpy} framework does not yet allow for full
    interoperability with \astropy. A goal of \astropy, and the scientific
    \python community is to enhance and implement the frameworks, allowing
    \astropyQuantity\ to propagate units seamlessly across the whole ecosystem of
    \package{numpy}-like projects.

>>>>

  % MHvK: this repeats the same thing 3 times; try to use just one sentence!
  The modular nature of \astropypkg means that not all sub-packages deeply
  integrate with one another -- and similarly between separate \python
  libraries. Because \astropy is situated at the nexus between scientific
  computing into astronomy, interoperability has been and continues to be an
  important focus for \astropy, both within the various \astropypkg sub-packages
  and with other important \python libraries. While increasing inter and
  intra-package integration will be a perpetual goal -- with feature additions
  in one place requiring updates in another -- \astropy has made impactful
  progress with units, tables, and I/O.
  % Better list? Also better transition
  If done right, all a user notices is that code using many different python
  features and libraries ``just works''.  Here, we describe a few
  particular efforts towards this goal.

@@

    \astropyQuantity\ is a backbone of \astropy, leveraging the power of
    \package{numpy} and adding units. Previously, many \package{numpy} functions
    would strip a \astropyQuantity\ of its units (or fail outright), limiting
    \astropyQuantity's usefulness. Now, advances in \package{numpy} function
    overloading (see for example
    \href{https://numpy.org/neps/nep-0018-array-function-protocol.html}{NEP 18})
    mean \astropyQuantity\ works with almost all \package{numpy} (version 1.17+)
    functions.
    % MHvK: I don't think the following is true, so commented it out.
    % These advances have similarly benefitted many
    % \package{numpy}-dependent packages, like \package{scipy} and their sparse
    % matrices. And again \astropypkg benefits since more \package{scipy}
    % functions now work with \astropyQuantity.

    In the remaining gaps, the \package{numpy} and \astropyQuantity\
    interoperability efforts are ongoing. For some functions, such as in
    \href{https://numpy.org/doc/stable/user/basics.rec.html#module-numpy.lib.recfunctions}{\package{numpy}'s
    module} for manipulating structured arrays, compatibility only requires
    exending the existing \package{numpy}-\astropypkg bridge framework.
    Community interest, in the form of a Feature Request (or better yet Pull
    Request) would be sufficient to see this realized. For a few remaining functions,
    discussed further in the
    \href{https://docs.astropy.org/en/stable/known_issues.html#known-deficiencies}{\astropy
    documentation}, the \package{numpy} framework does not yet allow for full
    interoperability with \astropy. A goal of \astropy, and the scientific
    \python community is to enhance and implement the frameworks, allowing
    \astropyQuantity\ to propagate units seamlessly across the whole ecosystem of
    \package{numpy}-like projects.
